Sequential Composition for Relaxed Memory: Pomsets with Predicate Transformers

### 2021-07-02 paper/drf.tex
finished with c11
changed example in C11

@@

The result requires that locations are properly initialized.  We assume a
sufficient condition: that programs have the form
``$\aLoc_1\GETS\aVal_1\SEMI \cdots \aLoc_n\GETS\aVal_n\SEMI\aCmd$'' where
every location mentioned in $\aCmd$ is some $\aLoc_i$.

We make two further restrictions to simplify the exposition.  To simplify the
definition of \emph{happens-before}, we ban fences and \RMWs.  To simplify
the proof, we assume there are no local declarations of the form
$(\VAR\aLoc\SEMI \aCmd)$.

@@

\bookmark{Comparison to Java}
In our language, past races are always resolved at a stable point, as in
\ref{Co3}.  As another example, consider the following, which is disallowed
here, but allowed by Java \cite[Ex.~2]{Dolan:2018:BDR:3192366.3192421}.  We
include an SC fence here to mimic the behavior of volatiles in the JMM.
\begin{gather*}
  \taglabel{past}
  \begin{gathered}
    (x\GETS 1 \SEMI   y^\mRA \GETS 1)
    \PAR
    (x\GETS 2 \SEMI \FENCE^\mSC \SEMI \IF{y^\mRA}\THEN r\GETS x \SEMI s\GETS x\FI)
    \\[-2ex]
    \hbox{\begin{tikzinline}[node distance=1.2em]
        \event{wx1}{\DW{x}{1}}{}
        \event{wy1}{\DWRel{y}{1}}{right=of wx0}
        \sync{wx1}{wy1}
        \event{wx2}{\DW{x}{2}}{right=3em of wy1}
        \event{f}{\DFS{\mSC}}{right=of wx2}
        \sync{wx2}{f}
        \event{ry1}{\DR[\mRA]{y}{1}}{right=of f}
        \sync{f}{ry1}
        \event{rx1}{\DR{x}{1}}{right=3em of ry1}
        \event{rx2}{\DR{x}{2}}{right=of rx1}
        \sync{ry1}{rx1}
        \sync[out=15,in=165]{ry1}{rx2}
        \rf[out=10,in=170]{wy1}{ry1}
        \wk[out=10,in=170]{wx1}{wx2}
        \wk[out=-170,in=-10]{rx1}{wx2}
        \bgellipsesmaller[yellow!50]{(wy1)(f)}{}
      \end{tikzinline}}
  \end{gathered}
\end{gather*}
The highlighted events are $L$-stable.  The order from $(\DR{x}{1})$ to
$(\DW{x}{2})$ is required by fulfillment, causing the cycle.  If the fence is
removed, there would be no order from $(\DW{x}{2})$ to
$(\DRAcq{y}{1})$, the highlighted events would no longer be $L$-stable, and
the execution would be allowed.  This more relaxed notion of ``past'' is not
expressible using \citeauthor{Dolan:2018:BDR:3192366.3192421}'s
synchronization primitives.

>>>>

Idea: merging reads can only make a difference is if there is a race.
Let $\sempomca{i}{}$ be the semantic function generated by applying the
construction of \reflem{lem:po} to $\semmca{i}{}$.

@@

The result requires that locations are properly initialized.  We assume a
sufficient condition: that programs have the form
``$\aLoc_1\GETS\aVal_1\SEMI \cdots \aLoc_n\GETS\aVal_n\SEMI\aCmd$'' where
every location mentioned in $\aCmd$ is some $\aLoc_i$.  To simplify the
definition of \emph{happens-before}, we ban fences and \RMWs.
% To simplify
% the proof, we assume there are no local declarations of the form
% $(\VAR\aLoc\SEMI \aCmd)$.

@@

\bookmark{Comparison to Java}
In our language, past races are always resolved at a stable point, as in
\ref{Co3}.  As another example, consider the following, which is disallowed
here, but allowed by Java \cite[Ex.~2]{Dolan:2018:BDR:3192366.3192421}.  We
include an SC fence here to mimic the behavior of volatiles in the JMM.
\begin{gather*}
  \taglabel{past}
  \begin{gathered}
    (x\GETS 1 \SEMI   y^\mRA \GETS 1)
    \PAR
    (x\GETS 2 \SEMI \FENCE{\mSC} \SEMI \IF{y^\mRA}\THEN r\GETS x \SEMI s\GETS x\FI)
    \\[-2ex]
    \hbox{\begin{tikzinline}[node distance=1.2em]
        \event{wx1}{\DW{x}{1}}{}
        \event{wy1}{\DWRel{y}{1}}{right=of wx0}
        \sync{wx1}{wy1}
        \event{wx2}{\DW{x}{2}}{right=3em of wy1}
        \event{f}{\DFS{\mSC}}{right=of wx2}
        \sync{wx2}{f}
        \event{ry1}{\DR[\mRA]{y}{1}}{right=of f}
        \sync{f}{ry1}
        \event{rx1}{\DR{x}{1}}{right=3em of ry1}
        \event{rx2}{\DR{x}{2}}{right=of rx1}
        \sync{ry1}{rx1}
        \sync[out=15,in=165]{ry1}{rx2}
        \rf[out=10,in=170]{wy1}{ry1}
        \wk[out=10,in=170]{wx1}{wx2}
        \wk[out=-170,in=-10]{rx1}{wx2}
        \bgellipsesmaller[yellow!50]{(wy1)(f)}{}
      \end{tikzinline}}
  \end{gathered}
\end{gather*}
The highlighted events are $L$-stable.  The order from $(\DR{x}{1})$ to
$(\DW{x}{2})$ is required by fulfillment, causing the cycle.  If the fence is
removed, there would be no order from $(\DW{x}{2})$ to
$(\DRAcq{y}{1})$, the highlighted events would no longer be $L$-stable, and
the execution would be allowed.  This more relaxed notion of ``past'' is not
expressible using \citeauthor{Dolan:2018:BDR:3192366.3192421}'s
synchronization primitives.
